Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.

### 2020-10-13 ccsn.tex
add NLC comments

\author{A.~Torres-Forn\'e}
\affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}

@@

\begin{abstract}
  Core collapse supernovae are very important phenomena in the Universe and their expected gravitational
  waves emission is a promising tool to study the onset of the supernova explosion mechanism. The complexity
  of the gravitational wave signal due to matter effects and the large degrees of freedom of the phenomena
  makes the source parameter inference problem very challenging. In this paper, we have considered the
  proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
  whose frequency time evolution is linked to the radius and the mass of the proton neutron star through
  universal relations demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract
  from the gravitational wave data the frequency evolution of the gravity mode. We have used a set of 1D
  core collapse supernova simulations to build a model of evolution of the proto neutron star properties
  with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
  the radius and mass of the proto neutron star. We have then estimated the performance of the method by
  performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
  11 and 40 solar mass and different equation of state. We have shown that for Advanced LIGO and Advanced
  Virgo detectors at design sensitivity it will be possible to infer poroton neutron star properties for
  a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow
  to test distances of several hundreds of kiloparsec.

>>>>

\author{A.~Torres-Forn\'e}
\affiliation{Max Planck Institute for Gravitationalphysik (Albert Einstein Institute), D-14476 Potsdam-Golm, Germany}
\affiliation{Departamento de Astronom\'{\i }a y Astrof\'{\i }sica, Universitat de Val\`encia, E-46100 Burjassot, Val\`encia, Spain}

@@

\begin{abstract}
  Core collapse supernovae are very important phenomena in the Universe and their expected gravitational
  waves emission is a promising tool to study the onset of the supernova explosion mechanism. The complexity
  of the gravitational wave signal due to matter effects and the large degrees of freedom of the phenomena
  makes the source parameter inference problem very challenging. In this paper, we have considered the
  proto neutron star oscillation modes that carry out a large fraction of the gravitational wave signal and
  whose frequency time evolution is linked to the system properties through universal relations as
  demonstrated in ~\cite{Torres:2019b}. We have developed a simple algorithm to extract from the
  gravitational wave data the frequency evolution of the gravity mode. A set of 1D core collapse
  supernova simulations is used to build a model of evolution of the proto neutron star properties
  with the gravity mode frequency. The model is then used to infer the time evolution of a combinaison of
  the radius and mass of the proto neutron star. We have then estimated the performance of the method by
  performing simulations with 2D core collapse supernova waveforms covering a progenitor mass range between
  11 and 40 solar mass and different equations of state. We have shown that for Advanced LIGO and Advanced
  Virgo detectors at design sensitivity it will be possible to infer proton neutron star properties for
  a galactic source, while third generation detectors Einstein Telescope and Cosmic Explorer will allow
  to test distances of several hundreds of kiloparsec.

### 2020-10-13 conclusion.tex
add NLC comments

\section{Conclusion}
The algorithm presented in this paper is a first attemp to infer the time evolution of a
combinaison of the mass of the PNS and its radius based on the universal relations found
in PNS asteroseismology. More precisely, we have considered in this paper the ratio
$r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$
oscillation mode in the GW data. We have especially investigated the performance of the algorithm
in the case of an optimally oriented source detected in a singe GW detector. For Advanced LIGO
or Advanced Virgo, the ratio can be reconstructed for a source in the Galaxy. We have shown
that this is true for a wide range of progenitor mass and that the quality of the inference
mainly depends on the signal-to-noise ratio of the signal. For third generation of GW detectors such
as Einstein Telescope and Cosmic Explorer, the $\mbox{}^2g_2$ will be reconstructible for sources
at distances of several kpc. Cosmic Explorer in its stage 2 configuration is obtaining the best performance
for all waveforms considered here thanks to its excellent sensitivity in the \unit{100-1000}[Hz] range.
Among the three configuration of Einstein Telescope, ET-D is providing the best performance,
especially for the waveforms with the highest progenitor mass (25 $\Msol$ and 40 $\Msol$). Comparing
$d_r$ for ET-B and the other third generations projects, it seems that having a good sensitivity
below \unit{200}[Hz] is important for massive mass progenitor signals.

This study does not include the realistic case of operating within a network of detectors. We defer this
for a forcoming publication. The sources of GWs we have considered here are optimally oriented. The
reported distance at which we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an
upper limit that may be lower by a factor 2--3 on average for a source located anywhere on the sky.

Finally, this method can be adapted to other PNS oscillation modes, changing few parameters such as
the frequency range of the beginning of the mode and its monotonic raise or descent. Being able to
reconstruct several modes in a same GW signal would allow to infer individually each of the PNS property.

>>>>

\section{Conclusion}
The algorithm presented in this paper is a first attempt to infer the time evolution of a
combinaison of the mass of the PNS and its radius based on the universal relations found
in PNS asteroseismology. More precisely, we have considered in this paper the ratio
$r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$
oscillation mode in the GW data. We have especially investigated the performance of the algorithm
in the case of an optimally oriented source detected in a singe GW detector. For Advanced LIGO
or Advanced Virgo, the ratio can be reconstructed for a source in the Galaxy. We have shown
that this is true for a wide range of progenitor masses and that the quality of the inference
mainly depends on the signal-to-noise ratio of the signal. For third generation of GW detectors such
as Einstein Telescope and Cosmic Explorer, the $\mbox{}^2g_2$ will be reconstructible for sources
at distances of several kpc. Cosmic Explorer in its stage 2 configuration is obtaining the best performance
for all waveforms considered here thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.
Among the three configuration of Einstein Telescope, ET-D is providing the best performance,
especially for the waveforms with the highest progenitor mass (25 $\Msol$ and 40 $\Msol$). Comparing
$d_r$ for ET-B and the other third generations projects, it seems that having a good sensitivity
below \unit{200}[Hz] is important for massive mass progenitor signals.

%The model to infer the ratio as function of the frequency is based on 1D {\sc AENUS-ALCAR} simulations.
%We have observed that using the other numerical code {\sc CoCoNut} inputs lead to systematic bias of the
%ratio reconstruction.

This study does not include the realistic case of operating within a network of detectors. We defer this
for a forthcoming publication. The sources of GWs we have considered here are optimally oriented. The
reported distance at which we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an
upper limit that may be lower by a factor 2--3 on average for a source located anywhere on the sky.

Finally, this method can be adapted to other PNS oscillation modes, changing few parameters such as
the frequency range of the beginning of the mode and its monotonic raise or descent. Being able to
reconstruct several modes in the same GW signal would allow to infer individually each of the PNS property.

### 2020-10-13 introduction.tex
add NLC comments

The life of massive stars {(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion. {Nearby} core-collapse supernova (CCSN) explosions {are expected to be sources of gravitational waves (GWs) and are one of the main candidates for the next great discovery by current ground-based observatories. However, these are relative rare events. A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slow rotating cores, which are present in the bulk of CCSN progenitors.
\mab{The emitted GWs} could be detected with advanced ground-based GW detectors \mab{(
  Advanced LIGO\citep{TheLIGOScientific:2014jea},
  Advanced Virgo\citep{TheVirgo:2014hva} and
  KAGRA\citep{Aso:2013eba})} within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful \mab{GW} signal {that could} be detected {within $50$~kpc and for some extreme models 
up to $5-30$ Mpc \cite{Gossan:2016,TargetedSNSearchO12}. However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated to long GRBs 
\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{add ref?}. Therefore, we focus this work only in neutrino-driven CCSNe. 
Despite the low rates, CCSN are of great scientific interest because they produce} a complex GW signals which could provide significant clues about the physical processes that occur in the moments after the collapse.

In the last decade, significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
% {[PCD: I don't think the next sentence is necessary since we have already discarded magneto-rotational events for this work.]
%\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood. The core-bounce signal can be directly related with the rotational properties of the core} \cite{Dimmelmeier:2007, abdikamalov:14, Richers:2017}. \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.  This} dynamics excite the different modes of oscillation of the PNS{, which ultimately leads to the emission of GWs. The frequency and time evolution of these modes carry information about the
properties of the GW emitter and could allow to perform PNS asteroseismology. }

@@

{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow {\mab oscillation} in the post bounce evolution with raising frequency
from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s. This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \cite{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}. In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers.

 The {properties of} g-modes in hot {PNSs} have been studied since the end of last century {by means of linear perturbation analysis of background PNS models}. The oscillation modes {associated to} the surface of hot PNSs was first considered by McDermott, van Horn \& Scholl \cite{McDermott:1983}. Additionally, the stratified structure of the PNS allows the presence of different types of g-modes related with the fluid core \cite{Reisenegger:1992}. Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \cite{Ferrari:2004}, general relativity \cite{Passamonti:2005}, non-linearities \cite{Dimmelmeier:2006}, phase transition \cite{Kruger:2015} and realistic equation of state \cite{Camelio:2017}. {Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \cite{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.

 % {We base this work in the PNS mode analysis performed by \cite{Torres:2018, Torres:2019a}, which explored
\mab{The eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
using results from 2D CCSN numerical simulations as a background studied in \cite{Torres:2018, Torres:2019a}
show a good match of the mode frequencies computed and the features observed in the GW spectrum of the same
simulation (specially when space-time perturbations are included \citep{Torres:2019a}).}
This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs.  {\cite{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}. These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}. {Similar relations have been found by \cite{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}

{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by \cite{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

>>>>

The life of massive stars {(those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$)} ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion. {Nearby} core-collapse supernova (CCSN) explosions {are expected to be sources of gravitational waves (GWs) and are one of the main candidates for the next great discovery by current ground-based observatories. However, these are relative rare events. A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slow rotating cores, which are present in the bulk of CCSN progenitors.
\mab{The emitted GWs} could be detected with advanced ground-based GW detectors \mab{(
Advanced LIGO\citep{TheLIGOScientific:2014jea},
Advanced Virgo\citep{TheVirgo:2014hva} and
KAGRA\citep{Aso:2013eba})} within $5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
For the case of fast rotating progenitor cores the result is likely a magneto-rotational explosion, with } a more powerful \mab{GW} signal {that could} be detected {within $50$~kpc and for some extreme models 
up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}. However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated to long GRBs 
\citep{Chapman:2007}), making this possibility a subdominant channel of detection with an event rate of $\sim 10^{-4} \rm{yr}^{-1}$\mab{[add ref?]}. Therefore, we focus this work only in neutrino-driven CCSNe. 
Despite the low rates, CCSN are of great scientific interest because they produce} a complex GW signals which could provide significant clues about the physical processes that occur in the moments after the collapse.

In the last decade, significant progress has been made in the development of numerical codes, {in particular in the treatment of multidimensioal effects \citep{BMueller:2020}.}
% {[PCD: I don't think the next sentence is necessary since we have already discarded magneto-rotational events for this work.]
%\sout{The waveforms produced by the magneto-rotational mechanism in particular is well understood. The core-bounce signal can be directly related with the rotational properties of the core} \citep{Dimmelmeier:2007, abdikamalov:14, Richers:2017}. \sout{However, the low rate of this kind of events and its expected low amplitude in the slow-rotation case, will probably impede its detection.}}
In the case of  neutrino-driven explosions, the GW emission is {primarly induced by instabilities developed at the newly formed proto-neutron star (PNS) and by the non-spherical accreting flow of hot matter over its surface \citep{Kotake:2017}.  This} dynamics excite the different modes of oscillation of the PNS{, which ultimately leads to the emission of GWs. The frequency and time evolution of these modes carry information about the
properties of the GW emitter and could allow to perform PNS asteroseismology. }

@@

{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} in the post bounce evolution with raising frequency
from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s. This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}. In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS.} The {typical} frequencies of {these} modes make them {a} promising {source} for ground-based interferometers.

 The {properties of} g-modes in hot {PNSs} have been studied since the end of last century {by means of linear perturbation analysis of background PNS models}. The oscillation modes {associated to} the surface of hot PNSs were first considered by McDermott {\it et al.} \citep{McDermott:1983}. Additionally, the stratified structure of the PNS allows for the presence of different types of g-modes related to the fluid core \citep{Reisenegger:1992}. Many posterior works used simplified neutron star models assuming an equilibrium configuration {as a background}, to study the effect of rotation \citep{Ferrari:2004}, general relativity \citep{Passamonti:2005}, non-linearities \citep{Dimmelmeier:2006}, phase transition \citep{Kruger:2015} and realistic equation of state \citep{Camelio:2017}. {Only recently, there has been an effort to incorporate realistic backgrounds based in numerical simulations in the computation of the mode structure and evolution \citep{Sotani:2016,Torres:2018, Morozova:2018, Torres:2019a,Torres:2019b,Sotani:2019,WS:2019,Sotani:2020a, Sotani:2020b}}.

 % {We base this work in the PNS mode analysis performed by \citep{Torres:2018, Torres:2019a}, which explored
\mab{The eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
using results from 2D CCSN numerical simulations as a background studied in \citep{Torres:2018, Torres:2019a}
shows a good match to the mode frequencies computed and the features observed in the GW spectrum of the same
simulation (specially when space-time perturbations are included \citep{Torres:2019a}).}
This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs. Authors in {\citep{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}. These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}. {Similar relations have been found by \citep{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}

{In this work, we present a method to infer from the GW data alone, the time evolution of some properties of the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \citep{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

### 2020-10-13 methods.tex
Update
update with final results
add NLC comments

In this section, we outline a strategy for estimating the time evolution of the
ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
oscillation mode in the gravitational wave detector data.
An integral part of this strategy is the universal relations that relate the
characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
and the radius of the PNS, the shock radius and the total mass inside the shock as
demonstrated in \cite{Torres:2019b}.

\pcd{To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
spherically symetric (1D) simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
shows the data for the $25$ numerical simulations. As identified by \cite{Torres:2019b}, the only systematic
deviation from a single universal relation is the numerical code used in the simulations. 
To avoid any systematic effect, we only use the $18$ simulations performed with the ALCAR-AENUS
code, which is the same code that was used in our test set. The consequences of this choice are discussed in the
conclusions. Using this data, we} parametrize the ratio with a cubic polynomial
regression with heteroscedastic errors
%Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polyn

@@

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}.} \label{fig:LMVAR}
\end{figure}

\pcd{We use this model to infer the properties of the simulations in the 
{\it test set} described in in Section \ref{sec:simulations}.}
To describe the method we focus on the gravitational wave signal
of {\texttt s20S}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates \pcd{[CITATION?]}of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes. 
\pcd{This is a property of the $\mbox{}^2 g_2$-mode, the frequency of which  
increases as the object becomes more massive and compact}.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/spectrogram}
 \caption{Spectrogram of the gravitational wave signal {\texttt s20S} sampled at \unit[4096]{Hz}.
   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
   with each other.} \label{fig:spectrogram}
\end{figure}

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use \pcd{our model given by Eq.~}
\eqref{eq:model1} to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. Ratio values
computed using the mass and radius values obtained from the simulation code \pcd{(true values)} are shown in black.

@@

In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\% \pcd{of the true values}.
In the next section we investigate the performance of \pcd{the} reconstruction of $r(t)$ when the gravitational wave
signal is embedded in noise.

>>>>

In this section, we outline a strategy for estimating the time evolution of the
ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
oscillation mode in the GW detector data.
An integral part of this strategy is the universal relations that relate the
characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
and the radius of the PNS, the shock radius and the total mass inside the shock as
demonstrated in \cite{Torres:2019b}.

To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
spherically symetric (1D) simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
shows the data for the $25$ numerical simulations. As identified by \cite{Torres:2019b}, the only systematic
deviation from a single universal relation is the numerical code used in the simulations. 
To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
code, which is the same code that was used in our test set.
%The consequences of this choice are discussed in the conclusions.
Using this data, we parametrize the discretized ratio $r_i$ with a cubic polynomial
regression with heteroscedastic errors
%Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polyn

@@

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from 25 1D simulations {\sc AENUS-ALCAR} (red) and {\sc CoCoNuT} (green) code. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering \pcd{only} the  {\sc AENUS-ALCAR} data points \pcd{($18$ simulations)}. \mab{To be redone in python}} \label{fig:LMVAR}
\end{figure}

{We use this model to infer the properties of the simulations in the 
{\it test set} described in in Section \ref{sec:simulations}.}
To describe the method we focus on the GW signal
of {\texttt s20S}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates \pcd{[add ref?]}of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t) $in the spectrogram,
taking into account that it is monotonically increasing as time goes. 
{This is a property of the $\mbox{}^2 g_2$-mode, the frequency of which  
increases as the object becomes more massive and compact}.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram}
 \caption{Spectrogram of the GW signal {\texttt s20S} sampled at \unit[4096]{Hz}.
   The spectrogram is obtained using data streach of 200 samples overlapping at 90\%
   with each other. \mab{To be redone in python}} \label{fig:spectrogram}
\end{figure}

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use our model given by Eq.~\eqref{eq:model1} to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An example is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. Ratio values
computed using the mass and radius values obtained from the simulation code \pcd{(true values)}
are shown in black.
In this example of a GW signal without noise the coverage of our 95\% confidence band is 100\%
of the true values.
In the next section we investigate the performance of the reconstruction of $r(t)$ when the GW
signal is embedded in noise.

@@
